---
title: "Analysis of COVID-era Apple Mobility Data.Rmd"
author: "Muhammed Hamir"
date: "10/10/2021"
output: html_document
bibliography: references.bib
params:
  state: "Alabama"
  data: "data/raw_data/applemobilitytrends-2021-10-09.csv"
  seqdata:output/sequence_summary.txt 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
```

## Setup

```{r set-parameters}
state_to_analyze <- params$state

# sanitizing the state name for output files
states_no_spaces <- gsub(state_to_analyze,
                         pattern = " ",
                         replacement = "_")

# set the name of the global data file to use
global_file_to_process <- params$data

# extract core name for output file naming use
core_name <- tools::file_path_sans_ext(basename(global_file_to_process))
```

```{r load-packages-functions}
# load the packages needed
library("dplyr")
library("readr")
library("ggplot2")
library("tidyr")
library("lubridate")

# load functions needed
source("code/functions/subset_mobility_data_to_state.R")
source("code/functions/tally_state_cities_counties.R")
source("code/functions/convert_mobility_wide_to_long.R")
```

## Background and Objectives

This rmarkdown file will use `ggplot2` [@ggplot2] and`dplyr`[@dplyr] packages to analyze the apple maps mobility data from the time before COVID-19 pandemic through the initial shutdowns and the continuing changes over the course of the year.

The goal will be to create a subset to a particular state, tally up the number of cities and counties in that state with mobility data, and then to make several plots to visualize those data subsets.

The state that this document will analyze is `r state_to_analyze`.

## Data Processing

```{r subset-state}
# subset out to the chosen state using the variables set above
state_subset <- subset_mobility_data_to_state(
  input_file_name = global_file_to_process,
  state_to_subset = state_to_analyze)
```

```{r make-subset-long}
# convert the subsetted state data from wide to long
long_subset <- convert_to_long(
  input_subsetted_data = paste0("output/subsetted_states_wide/",
                        core_name,
                        "_",
                        states_no_spaces,
                        ".csv"))
```

```{r tallied-subset}
# Tally up the numbers of cities and counties in the state that
# have data available
tallied_subset <- tally_state_cities_counties(
  input_state_file = paste0("output/subsetted_states_wide/",
                            core_name,
                            "_",
                            states_no_spaces,
                            ".csv"))
```


## figures 
```{r plot-tallied}
# create a grouped bar plot of the tallied data to include in the
# knitted document and save that figure to the output
# figures directory as a png file
tallied_plot <- ggplot(data = tallied_subset,
                       aes(x = geo_type,
                           y = n,
                           fill = transportation_type)) +
  geom_col(position = position_dodge()) +
  labs(title = paste("Number of cities and/or counties in",
                     state_to_analyze,
                     "with mobility data"),
       x = "Type of area",
       y = "Number of areas")

ggsave(plot = tallied_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         states_no_spaces,
                         "_tallied_plot.png"))
tallied_plot
```

```{r plot-time series}
# create a timeseries plot of the mean
# with date on the x axis and mean relative mobility on the y axis
# figures directory as a png file
timeseries_plot <- long_subset %>%
  filter(transportation_type == "driving") %>%
  group_by(date) %>%
  summarise(mean_mobility = mean(relative_mobility)) %>%
  ggplot(aes(x = lubridate::ymd(date),
             y = mean_mobility)) +
  geom_line() +
  labs(title = paste("Statewide mean relative mobility driving levels in",
                     state_to_analyze,
                     "during COVID"),
       x = "Date",
       y = "Mean relative mobility")

ggsave(plot = timeseries_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         states_no_spaces,
                         "_timeseies_plot.png"))
timeseries_plot
```

```{r plot-tallied2}
# create a grouped bar plot of the tallied data to include in the
# knitted document and save that figure to the output
# figures directory as a png file
tallied_plot2 <- ggplot(data = long_subset,
                       aes(x = geo_type,
                           y = relative_mobility,
                           fill = transportation_type)) +
  geom_col(position = position_dodge()) +
  labs(title = paste("Number of cities and/or counties in",
                     state_to_analyze,
                     "with relative mobility"),
       x = "Type of area",
       y = "Relative mobility")

ggsave(plot = tallied_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         states_no_spaces,
                         "_long_plot.png"))
tallied_plot2
```

```{r plot-sequence-data, fig.height=12}
# read in the count of sequences by country ( 2 columns)
seqs_by_country <- readr::read_table(params$seqdata,
                                     col_names = c("count",
                                                   "country"))
# check that we have the right number of columns
stopifnot(ncol(seqs_by_country) == 2)

# make a sideways barplot, log scale x axis, with countries on the y axis
# and the counts of sequences in each on the x, sorted by count
seq_count_plot <- ggplot(data = seqs_by_country,
                         aes(x = count,
                             y = reorder(country, count))) +
  geom_col() +
  scale_x_log10() +
  labs(title = "Count of SARS-CoV-2 sequences per country",
       x = "Count of Sequences",
       y = "Country Name")

ggsave(plot = seq_count_plot,
       filename = "output/figures/seq_count_plot.png")

seq_count_plot
```

## Session info
```{r session info}
sessionInfo()
```

### Sources Cited
```{r}
`
